Detecting the impact of land cover change on observed rainfall.

### 2019-07-05 Manuscripts/PeerJRainfall_Landuse/Rainfall_Landuse.tex
changes requested by PeerJ submission figures

\affil[1]{Sydney Institute of Agriculture, The University of Sydney, NSW 2006}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/map_selreg} \caption{Selected study regions are highlighted by red rectangles in the main map (the red rectangle in the insert indicates the location of the main map). The types of tree cover in 2008 from the DLCD product is shown at the background. In site 1 (the QLD region), the tree cover is mostly sparse. In site 2 (the NSW/VIC region), many areas have open or close forest where the tree cover is denser.}\label{fig:selreg}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/bushfire_nswvic} \caption{Location of bushfires occurring in January 2003, in and around the NSW/VIC study region, as shown by the red pixels. The map shows the large area in the Kosciuszko national park that has been burned. Some locations in the southwest of the Australian Capital Territory (ACT) have also experienced intensive bushfires.}\label{fig:bushfire}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/tc_simple} \caption{The expected evolution of the land surface after trees have been removed in (left) the QLD region and (right) the NSW/VIC region.}\label{fig:figure3-tc-simple}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/Rainfall_resid} \caption{The deseasonalised and detrended rainfall over the 30 years period in (a) the QLD region and (b) the NSW/VIC region. The vertical red lines indicate the year of 2003, in which the studied land cover changes occurred. A change in the time series data is not obvious before and after the land cover changes.}\label{fig:ts-mean}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/tc_figs} \caption{The maps show significant changes in tree cover identified from the MOD44B data between 2003 and 2015 in the NSW/VIC region (right) and the Qld region (left). The amount of change was calculated as the difference in tree cover before and after the specified land cover intervention and it is shown as the percentage of the ground area. Green colour indicates an increase in tree cover, while red colour indicates a decrease in tree cover.}\label{fig:tctrend}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.7\linewidth]{figures/rsq_fullmodel} \caption{The distribution of the variance explained (adjusted $r^2$ ) by the full regression model}\label{fig:rsqmodel}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.7\linewidth]{figures/soi_explhistogram} \caption{The performance of the regression model if rainfall is only modelled by the climate drivers. It shows the percentage of rainfall variability that can be explained by the climate drivers for the Qld and NSW/VIC region}\label{fig:rsq}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/Cp_30yrs} \caption{The spatial distribution of significance of land cover step change variable in the GAM model predicting changes in rainfall in the both regions. The left graph relates to the Qld region, while the right hand plot reflects the NSW/VIC region. The outlines of relevant Australian states and the Murray Darling Basin are indicated in grey.  The p value reported is for the land cover variable in the model.}\label{fig:LCp}
\end{figure}

The land cover variable in the model (Equation \eqref{eq:model2}) aims to
identify a step change in the rainfall before and after the observed
change in land cover. The variable was mainly significant
(\(p \leq 0.05\)) for the rainfall estimates in some areas in NSW/VIC,
as shown in Figure \ref{fig:LCp} (right panel). However, the number of
pixels where the landcover variable was significant was much greater in
NSW/VIC compared to the number in the QLD region. There was also some
relationship between the areas of bushfires in Figure
\ref{fig:bushfire}. Only a very small area with a significant step
change due to the land cover changes was found in rainfall in the QLD
region (left panel)

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/step_new} \caption{Spatial distribution of the step trend test Z' statistics in the two study sites. The panel on the left is for the QLD region, while the panel of the right is for NSW/VIC. Warm colours (yellow, orange and red) are for positive Z' values which indicate decreasing rainfall trend due to the land surface change. Cold colours (light blue to blue) are for negative Z' values which indicate increasing rainfall trend. The deeper the colour, the more significant the statistic.}\label{fig:steptest30}
\end{figure}

In both regions there are areas of positive Z' values which imply a
decrease in rainfall (Figure \ref{fig:steptest30}), but this is stronger
in the NSW/VIC area than in the QLD area. Qualitatively the locations
where changes in tree cover occurred in the NSW/VIC area (Figure
\ref{fig:tctrend}) seem to agree with the patterns in the Z' scores.
However, a direct relationship would not necesssarily be expected as
movement of air masses could mean that actual changes of rainfall are
observed close by, but not necessarily exactly at areas with changes of
landcover. For the QLD region (left panel) there are only a few
significant Z' scores which matches the earlier significance in the
landcover variable in the full regresion. In the QLD region, only 0.2\%
of the pixels obtained a positive Z' score with p \(<\) 0.1. In the
NSW/VIC region 3.7\% of pixels have a positive Z' score with p \(<\)
0.1. In general it is only a small proportion of both study regions.

\begin{figure}
\includegraphics[width=0.7\linewidth]{figures/bt4_hist} \caption{Field significance results showing the distribution of percentage of significant Z' scores for the resampled series (blue bars), and the precentage significant Z' scores for the original series of rainfall years (red circle)}\label{fig:fieldsig}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/slats} \caption{Woody vegetation clearing rate in QLD for the two major bioregions in the study area. The data were obtained from SLATS (2017), and clearly indicate a sharp decrease in the clearing rate after 2005.}\label{fig:slat}
\end{figure}

>>>>

\affil[1]{School of Life and Environmental Sciences, Sydney Institute of
Agriculture, The University of Sydney, New South Wales, Australia.}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/Fig1} \caption{Selected study regions are highlighted by red rectangles in the main map (the red rectangle in the insert indicates the location of the main map). The types of tree cover in 2008 from the DLCD product is shown at the background. In site 1 (the QLD region), the tree cover is mostly sparse. In site 2 (the NSW/VIC region), many areas have open or close forest where the tree cover is denser.}\label{fig:selreg}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/Fig2} \caption{Location of bushfires occurring in January 2003, in and around the NSW/VIC study region, as shown by the red pixels. The map shows the large area in the Kosciuszko national park that has been burned. Some locations in the southwest of the Australian Capital Territory (ACT) have also experienced intensive bushfires.}\label{fig:bushfire}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.7\linewidth]{figures/Fig3} \caption{The expected evolution of the land surface after trees have been removed in (A) the QLD region and (B) the NSW/VIC region.}\label{fig:figure3-tc-simple}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/Fig4} \caption{The deseasonalised and detrended rainfall over the 30 years period in (a) the QLD region and (b) the NSW/VIC region. The vertical red lines indicate the year of 2003, in which the studied land cover changes occurred. A change in the time series data is not obvious before and after the land cover changes.}\label{fig:ts-mean}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/Fig5} \caption{The maps show significant changes in tree cover identified from the MOD44B data between 2003 and 2015 in the Qld region (A) and the NSW/VIC region (B). The amount of change was calculated as the difference in tree cover before and after the specified land cover intervention and it is shown as the percentage of the ground area. Green colour indicates an increase in tree cover, while red colour indicates a decrease in tree cover.}\label{fig:tctrend}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.7\linewidth]{figures/Fig6} \caption{The distribution of the variance explained (adjusted $r^2$ ) by the full regression model}\label{fig:rsqmodel}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.7\linewidth]{figures/Fig7} \caption{The performance of the regression model if rainfall is only modelled by the climate drivers. It shows the percentage of rainfall variability that can be explained by the climate drivers for the Qld (A) and NSW/VIC (B) region}\label{fig:rsq}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/Fig8} \caption{The spatial distribution of significance of land cover step change variable in the GAM model predicting changes in rainfall in the both regions, the Qld region (A) and the NSW/VIC region (B). The outlines of relevant Australian states and the Murray Darling Basin are indicated in grey. The p-value reported is for the land cover variable in the model.}\label{fig:LCp}
\end{figure}

The land cover variable in the model (Equation \eqref{eq:model2}) aims to
identify a step change in the rainfall before and after the observed
change in land cover. The variable was mainly significant
(\(p \leq 0.05\)) for the rainfall estimates in some areas in NSW/VIC,
as shown in Figure \ref{fig:LCp} (B). However, the number of pixels
where the landcover variable was significant was much greater in NSW/VIC
compared to the number in the QLD region (Figure \ref{fig:LCp} (A)).
There was also some relationship between the areas of bushfires in
Figure \ref{fig:bushfire}. Only a very small area with a significant
step change due to the land cover changes was found in rainfall in the
QLD region (A)

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/Fig9} \caption{Spatial distribution of the step trend test Z' statistics in the two study sites, the QLD region (A) and the NSW/VIC region (B). Warm colours (yellow, orange and red) are for positive Z' values which indicate decreasing rainfall trend due to the land surface change. Cold colours (light blue to blue) are for negative Z' values which indicate increasing rainfall trend. The deeper the colour, the more significant the statistic.}\label{fig:steptest30}
\end{figure}

In both regions there are areas of positive Z' values which imply a
decrease in rainfall (Figure \ref{fig:steptest30}), but this is stronger
in the NSW/VIC area (B) than in the QLD area (A). Qualitatively the
locations where changes in tree cover occurred in the NSW/VIC area
(Figure \ref{fig:tctrend} (B)) seem to agree with the patterns in the Z'
scores. However, a direct relationship would not necesssarily be
expected as movement of air masses could mean that actual changes of
rainfall are observed close by, but not necessarily exactly at areas
with changes of landcover. For the QLD region (A) there are only a few
significant Z' scores which matches the earlier significance in the
landcover variable in the full regresion. In the QLD region, only 0.2\%
of the pixels obtained a positive Z' score with p \(<\) 0.1. In the
NSW/VIC region (B) 3.7\% of pixels have a positive Z' score with p \(<\)
0.1. In general it is only a small proportion of both study regions.

\begin{figure}
\includegraphics[width=0.7\linewidth]{figures/Fig10} \caption{Field significance results showing the distribution of percentage of significant Z' scores for the resampled series (blue bars), and the precentage significant Z' scores for the original series of rainfall years (red circle)}\label{fig:fieldsig}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/Fig11} \caption{Woody vegetation clearing rate in QLD for the two major bioregions in the study area. The data were obtained from SLATS (2017), and clearly indicate a sharp decrease in the clearing rate after 2005.}\label{fig:slat}
\end{figure}

### 2019-07-05 Manuscripts/PeerJRainfall_Landuse/SupplementaryMaterialRainfall_Landuse.Rmd
changes requested by PeerJ submission figures

---
title: Detecting the impact of land cover change on observed rainfall.
preprint: false
author:
  - name: Chun X. Liang
    affiliation: 1
  - name: Floris F. van Ogtrop
    affiliation: 1
  - name: R. Willem Vervoort
    affiliation: 1
    corresponding: true
    email: willem.vervoort@sydney.edu.au
affiliation:
  - code: 1
    address: Sydney Institute of Agriculture, The University of Sydney, NSW 2006
abstract: >
  This is the supplementary data for the paper "Detecting the impact of land cover change on observed rainfall." Peerj article number 35847 
header-includes: >
 \usepackage{lipsum}
 \usepackage{textcomp, rotating}
 \usepackage[normalem]{ulem}
bibliography: reflist_chap3.bib
date: "`r Sys.Date()`"
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article # for using bookdown features like \@ref()
  rticles::peerj_article: default
---
```{r setup, warning=F, message=F, echo =F}
# root dir
knitr::opts_knit$set(root.dir =
      # "E:/cloudstor/chun/chun/manuscripts/PeerJRainfall_Landuse")
      "C:/users/rver4657/owncloud/chun/chun/manuscripts/PeerJRainfall_Landuse")
```

@@

```{r cor-rain-qld, fig.cap="Cross-correlation of six climate indices and rainfall in QLD study region. For the PDO analysis, 108-year rainfall data (1900 - 2008) are used. Otherwise, 36-year rainfall data are used. The correlation with NINO 3 is not shown as it is very similar to but weaker than for NINO 3.4.", out.width="90%", echo=F}
include_graphics("figures/cor_qld.png")
```

```{r cor-rain-nsw, fig.cap="Cross-correlation of six climate indices and rainfall in NSW/VIC study region.", out.width="90%", echo=F}
include_graphics("figures/cor_nswvic.png")
```

@@

```{r residuals, fig.cap ="The residual analysis of a sample pixel in the QLD region (top) and NSW/VIC region (bottom).", echo=F, eval=T, out.width="90%"}
# removed after review
include_graphics("figures/gam_check.png")
```

@@

```{r meandiff, fig.cap ="Boxplots of annual rainfall residuals (estimated based on Equation 2 before and after the land cover intervention during 1979 - 2015 in the study regions. On average, the after period has a significantly lower annual rainfall residual in NSW/VIC, but a significantly higher annual rainfall residual in the Qld study area", echo=F, eval=T, out.width="70%"}
# figure dropped after review
include_graphics("figures/ResidualBoxplotchange.png")
```

>>>>

---
title: Detecting the impact of land cover change on observed rainfall.
preprint: false
author:
  - name: Chun X. Liang
    affiliation: 1
  - name: Floris F. van Ogtrop
    affiliation: 1
  - name: R. Willem Vervoort
    affiliation: 1
    corresponding: true
    email: willem.vervoort@sydney.edu.au
affiliation:
  - code: 1
    address: Sydney Institute of Agriculture, The University of Sydney, NSW 2006
abstract: >
  This is the supplementary data for the paper "Detecting the impact of land cover change on observed rainfall." Peerj article number 35847 
header-includes: >
 \usepackage{lipsum}
 \usepackage{textcomp, rotating}
 \usepackage[normalem]{ulem}
bibliography: reflist_chap3.bib
date: "`r Sys.Date()`"
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article # for using bookdown features like \@ref()
  rticles::peerj_article: default
---
```{r setup, warning=F, message=F, echo =F}
# root dir
knitr::opts_knit$set(root.dir =
      # "E:/cloudstor/chun/chun/manuscripts/PeerJRainfall_Landuse")
      "C:/users/rver4657/owncloud/chun/chun/manuscripts/PeerJRainfall_Landuse")
knitr::opts_knit$set('figcap.prefix','Figure S')
knitr::opts_knit$set('tabcap.prefix','Table S')
```

@@

```{r cor-rain-qld, fig.cap="Cross-correlation of six climate indices and rainfall in QLD study region. Bars in the plot indicate the strength of the cross-correlation at different lags. For the PDO analysis, 108-year rainfall data (1900 - 2008) are used. Otherwise, 36-year rainfall data are used. The correlation with NINO 3 is not shown as it is very similar to but weaker than for NINO 3.4.", out.width="90%", echo=F}
include_graphics("figures/FigS1.png")
```

```{r cor-rain-nsw, fig.cap="Cross-correlation of six climate indices and rainfall in NSW/VIC study region. Bars in the plot indicate the strength of the cross-correlation at different lags. For the PDO analysis, 108-year rainfall data (1900 - 2008) are used. Otherwise, 36-year rainfall data are used. The correlation with NINO 3 is not shown as it is very similar to but weaker than for NINO 3.4.", out.width="90%", echo=F}
include_graphics("figures/FigS2.png")
```

@@

```{r residuals, fig.cap ="The residual analysis of the Generalised additive modelling  indicating how well the residuals follow the regression assumptions. Results are shown for a sample pixel in the QLD region (top) and NSW/VIC region (bottom). Residuals are fairly normal as seen from the histogram and the qqplot, but show some scattering in the variance.", echo=F, eval=T, out.width="90%"}
# removed after review
include_graphics("figures/FigS3.png")
```

@@

```{r meandiff, fig.cap ="Boxplots of annual rainfall residuals (estimated based on Equation 2 before and after the land cover intervention during 1979 - 2015 in the study regions. On average, the after period has a significantly lower annual rainfall residual in NSW/VIC, but a significantly higher annual rainfall residual in the Qld study area", echo=F, eval=T, out.width="70%"}
# figure dropped after review
include_graphics("figures/FigS4.png")
```

### 2019-07-05 Manuscripts/PeerJRainfall_Landuse/SupplementaryMaterialRainfall_Landuse.tex
changes requested by PeerJ submission figures

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/cor_qld} \caption{Cross-correlation of six climate indices and rainfall in QLD study region. For the PDO analysis, 108-year rainfall data (1900 - 2008) are used. Otherwise, 36-year rainfall data are used. The correlation with NINO 3 is not shown as it is very similar to but weaker than for NINO 3.4.}\label{fig:cor-rain-qld}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/cor_nswvic} \caption{Cross-correlation of six climate indices and rainfall in NSW/VIC study region.}\label{fig:cor-rain-nsw}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/gam_check} \caption{The residual analysis of a sample pixel in the QLD region (top) and NSW/VIC region (bottom).}\label{fig:residuals}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.7\linewidth]{figures/ResidualBoxplotchange} \caption{Boxplots of annual rainfall residuals (estimated based on Equation 2 before and after the land cover intervention during 1979 - 2015 in the study regions. On average, the after period has a significantly lower annual rainfall residual in NSW/VIC, but a significantly higher annual rainfall residual in the Qld study area}\label{fig:meandiff}
\end{figure}

>>>>

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/FigS1} \caption{Cross-correlation of six climate indices and rainfall in QLD study region. Bars in the plot indicate the strength of the cross-correlation at different lags. For the PDO analysis, 108-year rainfall data (1900 - 2008) are used. Otherwise, 36-year rainfall data are used. The correlation with NINO 3 is not shown as it is very similar to but weaker than for NINO 3.4.}\label{fig:cor-rain-qld}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/FigS2} \caption{Cross-correlation of six climate indices and rainfall in NSW/VIC study region. Bars in the plot indicate the strength of the cross-correlation at different lags. For the PDO analysis, 108-year rainfall data (1900 - 2008) are used. Otherwise, 36-year rainfall data are used. The correlation with NINO 3 is not shown as it is very similar to but weaker than for NINO 3.4.}\label{fig:cor-rain-nsw}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figures/FigS3} \caption{The residual analysis of the Generalised additive modelling  indicating how well the residuals follow the regression assumptions. Results are shown for a sample pixel in the QLD region (top) and NSW/VIC region (bottom). Residuals are fairly normal as seen from the histogram and the qqplot, but show some scattering in the variance.}\label{fig:residuals}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.7\linewidth]{figures/FigS4} \caption{Boxplots of annual rainfall residuals (estimated based on Equation 2 before and after the land cover intervention during 1979 - 2015 in the study regions. On average, the after period has a significantly lower annual rainfall residual in NSW/VIC, but a significantly higher annual rainfall residual in the Qld study area}\label{fig:meandiff}
\end{figure}
